Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-09-01 paper.Rnw
adjusted figure size to meet submission requirements

# Specify global options
knitr::opts_chunk$set(echo = FALSE,
                      fig.keep = 'high',
                      message = FALSE)

@@

\begin{figure*}[!thp]
<<concept-plot-good, out.width = '\\textwidth', fig.width = 7.95, fig.height = 3.75, warning = FALSE>>=

@@

<<concept-plot-bad, out.width = '\\textwidth', fig.width = 7.95, fig.height = 3.75, warning = FALSE>>=

@@

\begin{figure*}[!thp]
\centering
<<sine-plot-data, out.width = '\\textwidth', fig.width = 8, fig.height = 3.1>>=

@@

\begin{figure*}[!thp]
<<sine-exp-scatter, out.width = '\\textwidth', fig.width = 12, fig.height = 4.68, warning = FALSE>>=

@@

\begin{figure}[!thp]
<<concept-feat-heat, out.width = '0.48\\textwidth', fig.width = 6, fig.height = 3>>=
# Specify the number of cases and LIME input options
ncases = 10
ninputs = 5

@@

<<sine-feat-heat, out.width = '0.45\\textwidth', fig.width = 4, fig.height = 4.5>>=
plot_feature_heatmap(sine_lime_explain$explain,
                order_method = "PCA") +
  theme_bw(base_family = "Times", base_size = 10) +
  theme(legend.position = "bottom",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_rect(color = "white", fill = "white")) +
  labs(y = "Case")
@
\caption{\emph{Feature Heatmap}. An example feature heatmap of the explanations from the applications of LIME with different tuning parameter values to the \data \ test set. The cases from the test set are plotted on the y-axis, and the tuning parameter values (simulation methods here) are included on the x-axis. The colors of the tiles indicate the feature selected by LIME for the corresponding case and tuning parameter value. The plot is faceted by the first and second most important features selected by LIME (top and bottom facets, respectively). The vertical facets separate the  kernel density method from the quantile bin methods. The vertical striping indicates that the LIME explanations are not consistent across tuning parameter values.}
\label{fig:sine-feat-heat}
\end{figure}

@@

\begin{figure}[!thp]
<<sine-assess-metric, out.width = '0.49\\textwidth', fig.width = 4.5, fig.height = 3.75, warning = FALSE>>=

@@

\begin{figure}[!t]
\centering
\includegraphics[width=2in]{./figure-static/bullets.png}
\caption{(Top left) Traditionally rifled gun barrel. The grooves and lands alternate to give bullets a spin during the firing process, which create markings (striations) on a bullet when fired. (Top right) Image of a fired bullet. The vertical stripes along the lower half of the bullet show groove and land engraved areas. The land engraved areas contain the microscopic striations created when the bullet passed through the barrel of the gun. (Bottom) Close up of a land engraved area showing striations (vertical lines).}
\label{fig:bullet}

@@

\includegraphics[width=3.25in]{./figure-static/signatures.png}
\caption{Example bullet signatures. The bullet signatures correspond to the same land and therefore have very similar patterns. The \citet{hare:2016} random forest is fit using various features that measure the similarity between two such signatures.}
\label{fig:signatures}
\end{figure}

@@

\begin{figure*}[!thp]
<<bullet-plot-rfpcp, out.width = '\\textwidth', warning = FALSE>>=

@@

\begin{figure*}[!thp]
<<bullet-feature-heatmap, out.width = '\\textwidth', fig.width = 18, fig.height = 12>>=
# Create a feature heatmap
plot_feature_heatmap(
  bullet_explain %>%
    mutate(
      label = as.factor(label),
      feature = fct_recode(
        feature,
        "Rough Correlation" = "rough_cor",
        "Consecutively Matching Striae" = "cms",
        "Distance" = "D",
        "Matches" = "matches",
        "Mismatches" = "mismatches",
        "Non-Consecutively Matching Striae" = "non_cms",
        "Cross Correlation Function" = "ccf",
        "Sum of Peaks" = "sum_peaks",
        "Distance Standard Deviation" = "sd_D"
      )
    ),
  facet_var = bullet_test %>%
    mutate(samesource = fct_recode(
      factor(samesource),
      "Match" = "TRUE",
      "Non-Match" = "FALSE"
    )) %>%
    pull(samesource) %>%
    na.omit(),
  order_method = "PCA"
) +
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) +
  scale_color_gretchenalbrecht(palette = "last_rays",
                               discrete = TRUE,
                               reverse = TRUE) +
  theme_bw(base_family = "Times", base_size = 16) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.y.right = element_text(angle = 0),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(y = "Case", color = "Complex Model Feature", fill = "Complex Model Feature")

@@

<<bullet-assess-metric, out.width = '\\textwidth', warning = FALSE, fig.width = 10, fig.height = 4>>=

@@

\begin{figure*}[!thp]
<<bullet-eois, out.width = '\\textwidth', fig.width = 12, fig.height = 17, fig.align = "center", out.width="90%", message = FALSE>>=
# Join the plots
plot_grid(
  pf2_3qb,
  pf2_3eb,
  eoi2_3qb,
  eoi2_3eb,
  pf1_3qb,
  pf1_3eb,
  eoi1_3qb,
  eoi1_3eb,
  nrow = 4,
  rel_heights = c(0.2, 0.3, 0.2, 0.3)
)
@
\caption{Plots of LIME explanations (first and third rows) and explanation scatterplots (second and fourth rows) for  cases M and NM in the bullet test data for two tuning parameter values: 3 quantile bins (first column) and 3 equal bins (second column).  The plots provide insights into the LIME explanations and allow for the assessment of the explanation quality.}
\label{fig:bullet-eois}
\end{figure*}

@@

\begin{figure*}[!thp]
<<sine-expl-scat-density, out.width = '\\textwidth', fig.width = 12, fig.height = 4.5, fig.align = "center">>=
plot_explain_scatter(
  sine_lime_explain$explain %>%
    filter(sim_method == "kernel_density", case == sine_poi %>% pull(case)),
  alpha = 0.75, title.opt = FALSE
) + 
  theme_bw(base_family = "Times") +
  theme(strip.background = element_rect(fill = "white", color = "white"),
        strip.placement = "outside")
@
\caption{Explanation scatterplots for the \data \ prediction of interest with the kernel density simulation method.}
\label{fig:sine-expl-scat-density}
\end{figure*}

\begin{figure*}[!thp]
<<bullet-expl-scat-density, out.width = '\\textwidth', fig.width = 12, fig.height = 9, fig.align = "center">>=
bullet_es_NM <-
  plot_explain_scatter(
  bullet_explain_perms[13:15, ] %>% mutate(case = "M"),
  alpha = 0.75, title.opt = TRUE
) +
  theme_bw(base_family = "Times") +
  theme(strip.background = element_rect(fill = "white", color = "white"),
        strip.placement = "outside")

@@

\begin{figure}[!thp]
<<sine-plot-resids, out.width = '0.5\\textwidth', fig.width = 5, fig.height = 4, warning = FALSE>>=

>>>>

# Specify global options
knitr::opts_chunk$set(
  echo = FALSE,
  fig.keep = 'high',
  message = FALSE,
  dev = "tiff",
  fig.align = 'center'
)

@@

\begin{figure*}[!thp]
<<concept-plot-good, out.width = '6.5in', fig.width = 7.95, fig.height = 3.75, warning = FALSE>>=

@@

<<concept-plot-bad, out.width = '6.5in', fig.width = 7.95, fig.height = 3.75, warning = FALSE>>=

@@

\begin{figure*}[!thp]
\centering
<<sine-plot-data, out.width = '6.5in', fig.width = 8, fig.height = 3.1>>=

@@

\begin{figure*}[!thp]
<<sine-exp-scatter, out.width = '6.5in', fig.width = 12, fig.height = 4.68, warning = FALSE>>=

@@

\begin{figure}[!thp]
<<concept-feat-heat, out.width = '3.125in', fig.width = 6, fig.height = 3>>=
# Specify the number of cases and LIME input options
ncases = 10
ninputs = 5

@@

<<sine-feat-heat, out.width = '3.125in', fig.width = 4, fig.height = 4.5>>=
plot_feature_heatmap(sine_lime_explain$explain,
                order_method = "PCA") +
  theme_bw(base_family = "Times", base_size = 10) +
  theme(legend.position = "bottom",
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.background = element_rect(color = "white", fill = "white")) +
  labs(y = "Case")
@
\caption{\emph{Feature Heatmap}. An example feature heatmap of the explanations from the applications of LIME with different tuning parameter values to the \data \ test set. The cases from the test set are plotted on the y-axis, and the tuning parameter values (simulation methods here) are included on the x-axis. The colors of the tiles indicate the feature selected by LIME for the corresponding case and tuning parameter value. The plot is faceted by the first and second most important features selected by LIME (top and bottom facets, respectively). The vertical facets separate the  kernel density method from the quantile bin methods. The vertical striping indicates that the LIME explanations are not consistent across tuning parameter values.}
\label{fig:sine-feat-heat}
\end{figure}

@@

\begin{figure}[!thp]
<<sine-assess-metric, out.width = '3.25in', fig.width = 4.5, fig.height = 3.75, warning = FALSE>>=

@@

\begin{figure}[!t]
\centering
\includegraphics[width=2.75in]{./figure-static/bullets.png}
\caption{(Top left) Traditionally rifled gun barrel. The grooves and lands alternate to give bullets a spin during the firing process, which create markings (striations) on a bullet when fired. (Top right) Image of a fired bullet. The vertical stripes along the lower half of the bullet show groove and land engraved areas. The land engraved areas contain the microscopic striations created when the bullet passed through the barrel of the gun. (Bottom) Close up of a land engraved area showing striations (vertical lines).}
\label{fig:bullet}

@@

\includegraphics[width=3.125in]{./figure-static/signatures.png}
\caption{Example bullet signatures. The bullet signatures correspond to the same land and therefore have very similar patterns. The \citet{hare:2016} random forest is fit using various features that measure the similarity between two such signatures.}
\label{fig:signatures}
\end{figure}

@@

\begin{figure*}[!thp]
<<bullet-plot-rfpcp, out.width = '6.5in', warning = FALSE>>=

@@

\begin{figure*}[!thp]
<<bullet-feature-heatmap, out.width = '6.5in', fig.width = 18, fig.height = 12>>=
# Create a feature heatmap
plot_feature_heatmap(
  bullet_explain %>%
    mutate(
      label = as.factor(label),
      feature = fct_recode(
        feature,
        "Rough Correlation" = "rough_cor",
        "Consecutively Matching Striae" = "cms",
        "Distance" = "D",
        "Matches" = "matches",
        "Mismatches" = "mismatches",
        "Non-Consecutively Matching Striae" = "non_cms",
        "Cross Correlation Function" = "ccf",
        "Sum of Peaks" = "sum_peaks",
        "Distance Standard Deviation" = "sd_D"
      )
    ),
  facet_var = bullet_test %>%
    mutate(samesource = fct_recode(
      factor(samesource),
      "Match" = "TRUE",
      "Non-Match" = "FALSE"
    )) %>%
    pull(samesource) %>%
    na.omit(),
  order_method = "PCA"
) +
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) +
  scale_color_gretchenalbrecht(palette = "last_rays",
                               discrete = TRUE,
                               reverse = TRUE) +
  theme_bw(base_family = "Times", base_size = 16) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.y.right = element_text(angle = 0),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(y = "Case", color = "Complex Model Feature", fill = "Complex Model Feature")

@@

<<bullet-assess-metric, out.width = '6.5in', warning = FALSE, fig.width = 10, fig.height = 4>>=

@@

\begin{figure*}[!thp]
<<bullet-eois, out.width = '6.5in', fig.width = 12, fig.height = 17, out.width="90%", message = FALSE>>=
# Join the plots
plot_grid(
  pf2_3qb,
  pf2_3eb,
  eoi2_3qb,
  eoi2_3eb,
  pf1_3qb,
  pf1_3eb,
  eoi1_3qb,
  eoi1_3eb,
  nrow = 4,
  rel_heights = c(0.2, 0.3, 0.2, 0.3)
)
@
\caption{Plots of LIME explanations (first and third rows) and explanation scatterplots (second and fourth rows) for  cases M and NM in the bullet test data for two tuning parameter values: 3 quantile bins (first column) and 3 equal bins (second column).  The plots provide insights into the LIME explanations and allow for the assessment of the explanation quality.}
\label{fig:bullet-eois}
\end{figure*}

@@

\begin{figure*}[!thp]
<<sine-expl-scat-density, out.width = '6.5in', fig.width = 12, fig.height = 4.5, fig.align = "center">>=
plot_explain_scatter(
  sine_lime_explain$explain %>%
    filter(sim_method == "kernel_density", case == sine_poi %>% pull(case)),
  alpha = 0.75, title.opt = FALSE
) + 
  theme_bw(base_family = "Times") +
  theme(strip.background = element_rect(fill = "white", color = "white"),
        strip.placement = "outside")
@
\caption{Explanation scatterplots for the \data \ prediction of interest with the kernel density simulation method.}
\label{fig:sine-expl-scat-density}
\end{figure*}

\begin{figure*}[!thp]
<<bullet-expl-scat-density, out.width = '6.5in', fig.width = 12, fig.height = 9, fig.align = "center">>=
bullet_es_NM <-
  plot_explain_scatter(
  bullet_explain_perms[13:15, ] %>% mutate(case = "M"),
  alpha = 0.75, title.opt = TRUE
) +
  theme_bw(base_family = "Times") +
  theme(strip.background = element_rect(fill = "white", color = "white"),
        strip.placement = "outside")

@@

\begin{figure}[!thp]
<<sine-plot-resids, out.width = '3.125in', fig.width = 5, fig.height = 4, warning = FALSE>>=
